Combining formal and distributional models of predicate semantics

### 2016-03-18 papers/acl2016/paper.tex
working on method
Former-commit-id: 76c883465605e8ee391de66afe27541fff462669

    \path[]
      (1) edge node [sloped, anchor=center, above] {\predicate{nationality}} (2)
      (1) edge node [sloped, anchor=center, above] {\predicate{type}} (3)
      (1) edge node [sloped, anchor=center, above] {\predicate{designed}} (5)
      (2) edge node [sloped, anchor=center, above] {\predicate{type}} (4)
      (5) edge node [sloped, anchor=center, above] {\predicate{located\_in}} (2);
  \end{tikzpicture}
  }
  \textbf{Features between \entity{Palladio} and \entity{Italy}}:\\
  \textless\predicate{nationality}\textgreater\\
  \textless\predicate{designed}, \predicate{located\_in}\textgreater
  \caption{A subset of the Freebase graph, and some example extracted features
  between \entity{Palladio} and \entity{Italy}.  The actual Freebase relations
  and entity identifiers used are modified here to aid readability.}
  \label{fig:sfe}
\end{figure}

@@

In this section we present our models of predicate semantics.  In
\secref{formal-and-distributional} we describe how we modify Krishnamurthy and
Mitchell's (K\&M's) semantic parser to include features from a formal knowledge
base.  In \secref{computing-pmi} we describe how we select those features for
each predicate.  The remaining two sections then describe other improvements we
made to K\&M's semantic parsing model: lexicalizing preposition and
noun-compound predicates, and using the Freebase graph to select candidate
entities.

\subsection{Formal features for predicate semantics}
\label{sec:formal-and-distributional}

We inject information from a formal knowledge base into our models of predicate
semantics simply by augmenting the learned distributional feature vectors with
formal features.

For each entity and entity pair in the data we compute an SFE feature vector
$\psi(e)$ and $\psi(e_1, e_2)$.  For entity pairs, the vector $\psi(e_1, e_2)$
is a sparse binary vector containing indicator features for all edge sequences
that connect nodes $e_1$ and $e_2$ in the Freebase graph, up to length 4.
These features are shown for the entity pair (\entity{Palladio},
\entity{Italy}) in the example graph in \figref{sfe}.  For entities, we extract
similar features to form the sparse binary vector $\psi(e)$.  These features
are of two types: (1) edge sequences that begin at node $e$ (up to length two),
and (2) edge sequences that begin at node $e$ along with the node at the end of
the sequence.  For example, all of the features listed in \figref{sfe} for the
entity pair (\entity{Palladio}, \entity{Italy}) are also valid features of type
(1) for the entity \entity{Palladio}, as well as the features
\textless\predicate{type}\textgreater and
\textless\predicate{designed}\textgreater.  Examples of of type (2) features
are \textless\predicate{type}\textgreater:\entity{architect}, and
\textless\predicate{nationality}\textgreater:\entity{Italy}.  Intuitively,
features of type (1) capture the fact that an entity participates in particular
kinds of relations, such as designing architectural structures, while features
of type (2) capture the way an entity is related to particular other entities.
Features of type (2) are very specific, and are most useful in practice when
the entity in the feature is a type node, such as with
\textless\predicate{type}\textgreater:\entity{architect}, though occasionally
features of this type involving other entities are also selected in our feature
selection process.

For each category and relation, we select a subset of the SFE features in the
vectors $\psi$ to have associated weights $\omega_c$ and $\omega_r$.  We then
add the weights $\omega_c$ and $\omega_r$ as parameters to be learned for each
predicate, and we add the feature vectors $\psi(e)$ and $\psi(e_1, e_2)$ as
observed features for each entity and entity pair.  The augmented category and
relation probabilities are as follows:

\begin{align*}
  P(c(e)) &= \sigma ( \theta_c^T \phi_e + \omega_c^T \psi_c(e)) \\
  P(r(e_1, e_2)) &= \sigma ( \theta_r^T \phi_{(e_1, e_2)} + \omega_r^T \psi_r(e_1, e_2) )
\end{align*}
where we have added subscripts on the $\psi$ functions to indicate that each
predicate selects its own set of features from the SFE feature vector $\psi$.

We pause here to note that these features obtained from SFE correspond exactly
to a limited set of horn clauses using the Freebase vocabulary, which in turn
can be seen as Freebase queries that define sets of entities or entity pairs.
By associating a number of these features with each predicate and assigning
them weights, what we are effectively doing is mapping language onto a weighted
combination of possible Freebase statements, instead of onto a single Freebase
query, as is typically done in the semantic parsing literature.  This is a
fundamentally new kind of predicate semantics that has not been used previously
in semantic parsers.

Note also that in our combined model there are now three sets of parameters to
be learned by the semantic parser: (1) $\theta$, low-dimensional distributional
vectors trained for each predicate; (2) $\phi$, low-dimensional distributional
vectors trained for each entity and entity pair; and (3) $\omega$, weights
associated with the selected formal SFE features for each predicate.  All of
these parameters are optimized jointly, using the same ranking objective used
by K\&M.

\subsection{Finding relevant features for each predicate}
\label{sec:computing-pmi}

The feature vectors returned by SFE have dimensionality in the tens of
millions.  We do not have near enough data to train that many parameters for
each of our tens of thousands of predicates, nor do we have enough RAM in our
machines to keep that many parameters in memory.  It is thus necessary to
select some small number of informative features for each predicate.  We do
this using pointwise mutual information (PMI).

PMI measures the strength of association between two variables $x$ and $y$; it
is computed as $\log\frac{p(x,y)}{p(x)p(y)}$.  In our case, we want to
associate predicates with SFE features.  However, our SFE features are for each
\emph{entity} and \emph{entity pair}, not each \emph{predicate}.  To associate
SFE features with predicates, we use the training data to find a mapping from
entities and entity pairs to predicates that they have been seen with in the
data.  For example, the phrase ``Italian architect Andrea Palladio'' is
considered a positive training example for the instantiated predicates
$\predicate{architect}(\entity{Andrea Palladio})$ and
$\predicate{architect\_N/N}(\entity{Italy}, \entity{Andrea Palladio})$.  We
then associate the feature vectors for \entity{Andrea Palladio} and
(\entity{Italy}, \entity{Andrea Palladio}) with the predicates
\predicate{architect} and \predicate{architect\_N/N}, respectively.  For each
predicate $\pi$ and feature $f$, we use these counts to calculate the
probabilities $p(\pi, f)$, $p(\pi)$ and $p(f)$.  After removing features with
counts below some threshold, we pick the $k$ features with the highest PMI
values for each predicate to use in our model.\footnote{We used $k=100$ in this
paper.}

>>>>

    \path[]
      (1) edge node [sloped, anchor=center, above] {\predicate{nationality}} (2)
      (1) edge node [sloped, anchor=center, above] {\predicate{type}} (3)
      (1) edge node [sloped, anchor=center, above] {\predicate{designed}} (5)
      (2) edge node [sloped, anchor=center, above] {\predicate{type}} (4)
      (5) edge node [sloped, anchor=center, above] {\predicate{located\_in}} (2);
  \end{tikzpicture}
  }
  \textbf{Features between \entity{Palladio} and \entity{Italy}}:\\
  \textless\predicate{nationality}\textgreater\\
  \textless\predicate{designed}, \predicate{located\_in}\textgreater

  \textbf{Features for \entity{Palladio}}:\\
  \textless\predicate{nationality}\textgreater\\
  \textless\predicate{nationality}\textgreater:\entity{Italy}\\
  \textless\predicate{type}\textgreater:\entity{Architect}\\
  \textless\predicate{designed}, \predicate{located\_in}\textgreater \\
  \textless\predicate{designed}, \predicate{located\_in}\textgreater:\entity{Italy}

  \caption{A subset of the Freebase graph, and some example extracted features
  between \entity{Palladio} and \entity{Italy}.  The actual Freebase relations
  and entity identifiers used are modified here to aid readability.}
  \label{fig:sfe}
\end{figure}

@@

This section presents our approach to open vocabulary semantic
parsing. Our approach follows the same structure as described in
Section \ref{sec:jayant-semparse}: we generate a surface logical form
from a query, then evaluate it against a probabilistic database with
learned predicate instance probabilities.  Section
\ref{sec:better-lfs} describes logical form generation, which improves
on prior work by lexicalizing additional predicates. Section

However, we improve both
logical form generation and the predicate models.

In
\secref{formal-and-distributional} we describe how we modify
Krishnamurthy and Mitchell's (K\&M's) semantic parser to include
features from a formal knowledge base.  In \secref{computing-pmi} we
describe how we select those features for each predicate.  The
remaining two sections then describe other improvements we made to
K\&M's semantic parsing model: lexicalizing preposition and
noun-compound predicates, and using the Freebase graph to select
candidate entities.

\subsection{Overview}
\label{sec:formal-and-distributional}

% TODO: ``query'' features?

We inject information from a knowledge base into our predicate models
by augmenting the learned embeddings with features. The truth
probability of a category instance $c(e)$ or relation instance $r(e_1,
e_2)$ in our model is given by:

\begin{align*}
  P(c(e)) &= \sigma ( \theta_c^T \phi_e + \omega_c^T \psi_c(e)) \\
  P(r(e_1, e_2)) &= \sigma ( \theta_r^T \phi_{(e_1, e_2)} + \omega_r^T \psi_r(e_1, e_2) )
\end{align*}

In the above equation, the $\theta$ and $\phi$ terms are learned
predicate and entity embeddings, as described in Section
\ref{sec:jayant-semparse}. The second term in the sum represents our
new features and their learned weights. $\psi_c(e)$ and $\psi_r(e_1,
e_2)$ are SFE feature vectors for each entity and entity pair; a
different set of features is chosen for each predicate in order to
maximize their usefulness. The terms $\omega_c$ and $\omega_r$
represent learned weights for these features.

Each SFE feature can be viewed as either a Freebase query or a horn
clause implication rule using the Freebase schema. In both cases, the
feature defines a set of entities (or entity pairs), and the
associated feature weight measures the likelihood that an entity in
this set is also in the denotation of the surface predicate. Our
models include \emph{many} such features for each surface predicate,
effectively mapping each surface predicate onto a \emph{weighted
  combination of Freebase queries}. This approach is fundamentally
different from typical semantic parsers that map language to a single
Freebase query.

In our model, there are now three sets of parameters to be learned:
(1) $\theta$, low-dimensional distributional vectors trained for each
predicate; (2) $\phi$, low-dimensional distributional vectors trained
for each entity and entity pair; and (3) $\omega$, weights associated
with the selected formal SFE features for each predicate.  All of
these parameters are optimized jointly, using the same ranking
objective used by K\&M (TODO: not sure how to eliminate this backref).

% For each category and relation, we select a subset of the SFE features in the
% vectors $\psi$ to have associated weights $\omega_c$ and $\omega_r$.  We then
% add the weights $\omega_c$ and $\omega_r$ as parameters to be learned for each
% predicate, and we add the feature vectors $\psi(e)$ and $\psi(e_1, e_2)$ as
% observed features for each entity and entity pair.  The augmented category and
% relation probabilities are as follows:

\subsection{Generating predicate features}

The feature vectors $\psi_c(e)$ and $\psi_r(e_1, e_2)$ are constructed
using a two-stage process. First, we generate a predicate-independent
feature vector $\psi(e)$ for each entity $e$ and $\psi(e_1, e_2)$ for
each entity pair $(e_1, e_2)$ using SFE. This step generates very
high-dimensional feature vectors, where most of the
features are not relevant for any given predicate. Next, we perform
per-predicate feature selection to identify a subset of these features
that are relevant for each category and relation predicate.

The predicate-independent feature extraction stage computes the
following feature vectors. For entity pairs, the vector $\psi(e_1,
e_2)$ is a sparse binary vector containing indicator features for all
edge sequences that connect nodes $e_1$ and $e_2$ in the Freebase
graph, up to length 4.  For an entity $e$, we extract two kinds of
features: (1) all edge sequences up to length two that begin at $e$,
and (2) the same sequences paired with the entity at their
endpoint. Features of type (1) capture the kinds of relations that an
entity participates in, e.g., designing architectural structures,
while features of type (2) capture relations to other particular
entities. This second kind of feature is most helpful when the
endpoint entity is a type, such as \entity{architect}, although there
are some other valuable features of this kind. \figref{sfe} shows some
example entity and entity pair features extracted from a subset of the
Freebase graph.

% For each category and relation, we select a subset of the SFE features in the
% vectors $\psi$ to have associated weights $\omega_c$ and $\omega_r$.  We then
% add the weights $\omega_c$ and $\omega_r$ as parameters to be learned for each
% predicate, and we add the feature vectors $\psi(e)$ and $\psi(e_1, e_2)$ as
% observed features for each entity and entity pair.  The augmented category and
% relation probabilities are as follows:
% 
% \begin{align*}
%   P(c(e)) &= \sigma ( \theta_c^T \phi_e + \omega_c^T \psi_c(e)) \\
%   P(r(e_1, e_2)) &= \sigma ( \theta_r^T \phi_{(e_1, e_2)} + \omega_r^T \psi_r(e_1, e_2) )
% \end{align*}
% where we have added subscripts on the $\psi$ functions to indicate that each
% predicate selects its own set of features from the SFE feature vector $\psi$.

% The feature vectors returned by SFE have dimensionality in the tens of
% millions.  We do not have near enough data to train that many parameters for
% each of our tens of thousands of predicates, nor do we have enough RAM in our
% machines to keep that many parameters in memory.  It is thus necessary to
% select some small number of informative features for each predicate.  We do
% this using pointwise mutual information (PMI).

The feature selection step uses pointwise mutual information (PMI) to
reduce the number of features considered by the model for each
predicate. The feature vectors generated above have tens of millions
of dimensions, but, for any given surface predicate, only a handful of
these dimensions represent relevant Freebase queries. We select
features by first summing the entity and entity pair feature vectors
seen with each predicate in the training data. For example, the phrase
``Italian architect Andrea Palladio'' is considered a positive
training example for the instantiated predicates
$\predicate{architect}(\entity{Andrea Palladio})$ and
$\predicate{architect\_N/N}(\entity{Italy}, \entity{Andrea
  Palladio})$. We then add the feature vectors for \entity{Andrea
  Palladio} and (\entity{Italy}, \entity{Andrea Palladio}) to the
feature counts for the predicates \predicate{architect} and
\predicate{architect\_N/N}, respectively. Next, for each predicate
$\pi$ and feature $f$, we use these counts to calculate three
probabilities: $p(\pi, f)$, $p(\pi)$ and $p(f)$. (TODO: how did you
calculate these?). The PMI of a predicate/feature pair is then $\log
\frac{p(\pi, f)}{p(\pi)p(f)}$. After removing features with counts
below a threshold, we pick the $k=100$ features with the highest PMI
values for each predicate to use in our model.

% PMI measures the strength of association between two variables $x$ and $y$; it
% is computed as $\log\frac{p(x,y)}{p(x)p(y)}$.  In our case, we want to
% associate predicates with SFE features.  However, our SFE features are for each
% \emph{entity} and \emph{entity pair}, not each \emph{predicate}.  To associate
% SFE features with predicates, we  For each
% predicate $\pi$ and feature $f$, we use these counts to calculate the
% probabilities $p(\pi, f)$, $p(\pi)$ and $p(f)$.
